\LARGE \bf TODO Name this thing

### 2012-02-17 papers/itsc_2012/root.tex
paper: architecture, map model, determinism properties

This paper presents a new methodology for micro-simulating autonomous traffic for such situations where we will continue to use existing roads for autonomous vehicles. We believe that the autonomous traffic will be controlled via an overseer. The overseer can define behaviors for indivudal agents and individual intersections. The overseer itself can be a centralized or decentralized system that is capable of making such decisions in real-time.

@@

AORTA's implementation uses Scala, a language implemented on the Java Virtual
Machine. Scala provides the advantage of functional programming constructs
without imposing them; parts of AORTA or extensions or clients can just as
easily be written in the more familiar Java.

The architecture of AORTA is described below.

thread pools?

programs: build map, UI (describe what it can do -- particularly demand
generation), headless

api / extending it
maps separate from simulation separate from UI (mostly)

@@

\commentd{briefly introduce OSM, advantages, etc}

@@

roads, verts, edges, wards, traversables, turns...

edges are made up of multiple lines.. this could be control pts of a curve
(hereon, line means line segment)
(turns are also traversable)

illustration of turn conflicts

@@

@@

We are also experimenting with explicitly stating where traffic should spawn and where
it should go.  This allows more accurate (and consistent) simulations of situations such
as rush hour.  In the future, we hope to be able to generate these demand descriptions
from government-collected data about a municipality's road usage.

@@

justify arbitrary constants, like deaccel for cars

scale \commentm{I'm going to change this back to a 1:1 correlation with the real world.  (Hopefully this weekend.)}

systems ideas, or limitations.. trylocks, atomic sim update, deadlock issues

limit: no lane-changing

@@

\commentd{architecture review (generators and thread pool)}

@@

TODO

@@

>>>>

\commentd{nevada even made autonomous cars legal:
http://www.pcmag.com/article2/0,2817,2400400,00.asp}
This paper presents a new methodology for micro-simulating autonomous traffic for such situations where we will continue to use existing roads for autonomous vehicles. We believe that the autonomous traffic will be controlled via an overseer. The overseer can define behaviors for indivudal agents and individual intersections. The overseer itself can be a centralized or decentralized system that is capable of making such decisions in real-time.

@@

\commentd{UML diagram?}

The architecture and implementation notes of AORTA follow. AORTA is divided into
roughly three components, each somewhat independent of the others. The map model
handles everything from building graphs from OSM source to pathfinding. This
component is cleanly separated from the simulation engine, which adds the notion
of agents, structures on top of map components that detect collisions, and the
machinery to implement AORTA's model of car dynamics. If a future experiment
demands macrosimulation or cellular automata method, no changes to the map
infrastructure are needed. Finally, there is a UI using Java Swing, which is
(out of necessity) tied to both the map and simulation components. The software
has a documented API, is open-source, and contains all of the code for the
experiments described in Section 6: it is designed for extensibility.

AORTA's implementation uses Scala, a language implemented on the Java Virtual
Machine. Scala provides the advantage of functional programming constructs while
still permitting imperative style. Parts of AORTA or extensions or clients can
just as easily be written in Java.

There are three standalone applications in AORTA. The map builder takes an OSM
file as input and produces an AORTA graph as output. There is a UI, using Swing,
to interactively explore the map and run simulations. Specific traffic patterns
can quickly be simulated by selecting a source and destination group of roads in
the UI. Finally, there is a headless mode that runs some specific experiment
without the overhead of visualization. The UI and headless mode both take a
graph as input, meaning the map construction step must be run once per OSM
input, but the same graph can be reused indefinitely.

@@

Our goal is to simulate traffic on existing infrastructure, not contrived,
randomly generated maps. We do this by utilizing OpenStreetMaps, a community
project offering user-editable, free maps.

@@

AORTA's map format reflects our map model, which will now be described.

\commentd{terminology sucks below. suggestions?}

\emph{Roads} are undirected edges between \emph{vertices}, which are
intersections between multiple roads. A road contains one or two sets of
directed \emph{edges}, each of which represents a single lane of travel. One-way
roads are supported. An edge leads from one vertex to another; the second vertex
then contains a mapping from each incoming edge to some outgoign edge; each pair
is a \emph{turn}. Strictly speaking, the multigraph consists of \emph{lanes} as
vertices (this is confusing, because lanes are \emph{edges} in our model) and
\emph{turns} as edges. \emph{Roads} are a super-structure with meta-data,
bundling several \emph{lanes} together, and \emph{vertices} serve to bundle
together \emph{turns}.

Geometrically speaking, a road has a sequence of points defining the line
segments constituting the center yellow line. These points could also be
interpreted as control points for a curved road, but AORTA approximates using
straight lines. Edges mimic this geometry, projecting the center line a lane's
width away. Turns are modeled with a single straight line connecting the end of
one edge to the start of another. Since both edges and turns can support
traffic, they are grouped together as \emph{traversables}, meaning they have
some sequence of line segments and a total length.

@@

\commentd{illustration of turn conflicts}

AORTA has a low-granularity model of turn conflict. If two vehicles at any point
along two different turns could ever come into contact, then the turns are
always in conflict, no matter where the agent specifically is along that turn.
The alternative of tiling intersections would provide more granularity, but it
adds computational complexity and would be impractical for many of the
intersection geometries inferred from OSM.

@@

Section 5.1 details a way to specify where traffic should originate and move
towards using the UI. This allows a more accurate simulation of situations such
as rush hour. In the future, we hope to be able to generate these demand
descriptions from government-collected data about a municipality's road usage.

@@

\commentd{justify arbitrary constants, like deaccel for cars;
          systems ideas, or limitations.. trylocks, atomic sim update, deadlock
          issues;
          limit: no lane-changing}

@@

\commentd{(generators and thread pool)}

@@

AORTA is designed for controlled experiments. Consequently, simulations are
deterministically reprodcible. A single seed for the pseudo-random number
generator, the input graph, and the generators completely determine the single
outcome of any simulation. This determinism breaks down when generators use
worker threads, since routes may be computed in different orders, causing agents
to be ready to enter the road at different times.

\commentd{and how will we solve this issue with threads and determinism?}

\commentd{give some informal or formal (with numbers) idea of scale / how many
agents we can simulate in different map sizes}

@@

What credits do we need to give to OSM?
